Pillar 1 - Sequential Pattern Mining

### 2019-01-17 Pillar 1 - Sequential Pattern Mining.Rmd

# Setup
```{r}
load("data_pillar_1.RDATA")
```

@@

results %>%
  arrange(desc(confidence))
```

```{r}
d_transcript %>% filter(`Course Code`=="SSC3016")
```

Now we include data for pass and fail
```{r}
d_transaction_passed <- d_course %>%
  select(Type, Code, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(Code) %>%
  left_join(d_transcript, by = c("Code" = "Course Code")) %>%
  mutate(Outcome = case_when(Grade=="NG"~"Fail",
                             Grade > 5.5 ~"Pass",
                             T~"Fail")) %>%
  mutate(Item= paste(Code, Outcome, sep = "_")) %>%
  select(`Student ID Number`, Item) %>%
  group_by(`Student ID Number`) %>%
  summarize(list_course = list(Item)) %>%
  group_by()

@@

```{r}
rules <- ruleInduction(frequent_trajectories, confidence = 0.2, #we need to play with the parameters.
                    control    = list(verbose = TRUE))
as(rules, "data.frame") %>%
  arrange(desc(lift))
```
#Sequence mining with pass fail.
First we add pass fai information to our data set and transform it to transactions class:
```{r}
#sequences
d_pass_sequence_transactions <- d_course %>%
  select(Type, Code, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(Code) %>%
  left_join(d_transcript, by = c("Code" = "Course Code")) %>%
  mutate(Outcome = 
           case_when( Grade=="NG"~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade > 5.5 ~"Pass",
                              T~"Fail")
         ) %>%
  mutate(Item = paste(Code, Outcome, sep = "_") ) %>%
  filter(!str_detect(Period, " to ")) %>% #filter courses that spann more than one period
  mutate( Time = paste(Year_numerical, Period, sep = ""),
          Time = as.double(Time)) %>%
  select(`Student ID Number`, Item, Time) %>%
  arrange(`Student ID Number`, Time) %>% #Note that integer identifiers must be positive and that transactions must be ordered by sequenceID and eventID.
  group_by(`Student ID Number`, Time) %>%
  summarize(list_course = list(Item)) %>% #if we run the code until here maybe its possible to use sequence_transactions <- as(sequence_transaction$list_course, "transactions")
  mutate(TransactionID = paste(`Student ID Number`, Time, sep = "_"))

@@

pass_rules <- ruleInduction(frequent_pass_trajectories, confidence = 0.2, #we need to play with the parameters.
                    control    = list(verbose = TRUE))
pass_rules_df <- as(pass_rules, "data.frame") %>%
  arrange(desc(lift))
```
Checking who failed
```{r}
pass_rules_df %>%
  separate(rule, into=c("LHS", "RHS"), sep = "=>", remove = F) %>%
  filter(str_detect(RHS, "_Fail"))
```

>>>>

# Setup
load previous data and create function to find courses
```{r}
load("data_pillar_1.RDATA")

find_course <- function(data = d_transcript, code){
  c = code
  x <- data %>%
    filter(`Course Code`== c)

  return(x$`Course Title`[1])
}

@@

results %>%
  arrange(desc(confidence))
```
Now we include data for pass and fail:
```{r}
d_transaction_passed <- d_course %>%
  select(Type, Code, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(Code) %>%
  left_join(d_transcript, by = c("Code" = "Course Code")) %>%
  mutate(Outcome = case_when(Grade=="NG"~"Fail",
                             Grade > 5.5 ~"Pass",
                             T~"Fail")) %>%
  mutate(Item= paste(Code, Outcome, sep = "_")) %>%
  select(`Student ID Number`, Item) %>%
  group_by(`Student ID Number`) %>%
  summarize(list_course = list(Item)) %>%
  group_by()

@@

```{r}
rules <- ruleInduction(frequent_trajectories, confidence = 0.2, #we need to play with the parameters.
                    control    = list(verbose = TRUE))
as(rules, "data.frame") %>%
  arrange(desc(lift))
```
#Sequence mining with pass fail.
First we add pass fai information to our data set and transform it to transactions class:
```{r}
#sequences
fail_threshold <- 6.5

d_pass_sequence_transactions <- d_course %>%
  select(Type, Code, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(Code) %>%
  left_join(d_transcript, by = c("Code" = "Course Code")) %>%
  mutate(Outcome = 
           case_when( Grade=="NG"~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade > 5.5 ~"Pass",
                              T~"Fail"),
         Low_grade =
           case_when(Grade=="NG"~"Sad", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"OK",
                              T~"Sad")
         ) %>%
  mutate(Item = paste(Code, Low_grade, sep = "_") ) %>% #Change for Outcome/Low_grade
  filter(!str_detect(Period, " to ")) %>% #filter courses that spann more than one period
  mutate( Time = paste(Year_numerical, Period, sep = ""),
          Time = as.double(Time)) %>%
  select(`Student ID Number`, Item, Time) %>%
  arrange(`Student ID Number`, Time) %>% #Note that integer identifiers must be positive and that transactions must be ordered by sequenceID and eventID.
  group_by(`Student ID Number`, Time) %>%
  summarize(list_course = list(Item)) %>% #if we run the code until here maybe its possible to use sequence_transactions <- as(sequence_transaction$list_course, "transactions")
  mutate(TransactionID = paste(`Student ID Number`, Time, sep = "_"))

@@

pass_rules <- ruleInduction(frequent_pass_trajectories, confidence = 0.1, #we need to play with the parameters.
                    control    = list(verbose = TRUE))
pass_rules_df <- as(pass_rules, "data.frame") %>%
  arrange(desc(lift))
```
Checking who failed
```{r}
pass_rules_df %>%
  separate(rule, into=c("LHS", "RHS"), sep = "=>", remove = T) %>% #remove= F to keep columns
  filter(str_detect(RHS, "_Sad")) %>%
  arrange(desc(confidence), desc(lift))
```
